---
title: "Homework_10"
author: "Claire Kraft"
date: "2024-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 14.1

**The breast cancer data set breast-cancer-wisconsin.data.txt from [UCI Repository](http://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/) (description at [UCI Dataset Description](http://archive.ics.uci.edu/ml/datasets/Breast+Cancer+Wisconsin+%28Original%29)) has missing values.**

1. **Use the mean/mode imputation method to impute values for the missing data.**

2. **Use regression to impute values for the missing data.**

3. **Use regression with perturbation to impute values for the missing data.**

4. **(Optional) Compare the results and quality of classification models (e.g., SVM, KNN) built using**  
   **(1) the data sets from questions 1, 2, 3;**  
   **(2) the data that remains after data points with missing values are removed;**  
   **(3) the data set when a binary variable is introduced to indicate missing values.**

  
Went to the description [URL](http://archive.ics.uci.edu/dataset/15/breast+cancer+wisconsin+original) to understand the schema of the table. Column 11 is "Class" and the datatype is binary. If the value is 2 then it's benign and if the value is 4 then it's malignant. According to the aggregation, there 241 malignant and 458 benign cases. There are 691 total missing data points in my dataset. The BareNuclei column has 2.2% of the data (with N of 16) missing.

There is a strong correlation between the cell size and shape.

```{r breast cancer}
# Library imports
library(tidyverse)
library(data.table)
#install.packages("visdat")
library(visdat)
library(ggplot2)
#install.packages("naniar")
library(naniar)
library(GGally) #ggplot2 extension for pairs matrix
#install.packages("psych")
library(psych) 
library(rpart)
#install.packages("rpart.plot")
library(rpart.plot)

# Read in data
df <- fread('https://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/breast-cancer-wisconsin.data', header=FALSE, stringsAsFactors = FALSE)
# Preview data
head(df)
summary(df)

# Count observations per class
df %>%
  count(V11) 
```

```{r breast cancer - cleaning: transform}
# Copy original df
df_cleaning <- df

# Rename columns
df_cleaning <- rename (df_cleaning, 
                       SampleCode=V1, 
                       ClumpThickness=V2, 
                       CellSize=V3, 
                       CellShape=V4, 
                       Adhesion=V5, 
                       SingleEpithelialSize=V6, 
                       BareNuclei=V7, 
                       BlandChromatin=V8, 
                       NormalNucleoli=V9,
                       Mitoses=V10, 
                       Diagnosis=V11)

# Convert BareNuclei or V7 column to integer
df_cleaning$BareNuclei <- as.integer(df_cleaning$BareNuclei)

# Convert Diagnosis or V11 to factor
df_cleaning$Diagnosis <- as.factor(df_cleaning$Diagnosis)

# Set the levels of the Diagnosis factor to 0 and 1. This effectively assigns new labels to the existing factor levels.
levels(df_cleaning$Diagnosis) <- c(0, 1)

summary(df_cleaning)
head(df_cleaning)
```

```{r breast cancer - cleaning: dupes}
# Dupes check
n_distinct(df_cleaning) # 691

# Zoom in on the dupes
duplicates <- df_cleaning %>%
  filter(duplicated(.))

print(duplicates)
```

```{r breast cancer - cleaning: missing}
missing_col_summary <- colSums(is.na(df_cleaning))
print(missing_col_summary)

# Get a summary of missing values
miss_var_summary(df_cleaning)

print(sprintf("Percent of missing observation = %0.3f", 16/nrow(df_cleaning)*100))

# Visualize missing values
vis_miss(df_cleaning)
```
### Imputation - mean/mode
```{r breast cancer - cleaning: mean imputation}
#---------------------- Mean imputation
df_cleaning.mean<-df_cleaning
df_cleaning.mean<-df_cleaning.mean %>% mutate_at(vars(BareNuclei),~ifelse(is.na(.x), mean(.x, na.rm = TRUE), .x))

# Double check mean imputation
head(df_cleaning.mean,24)
```
```{r breast cancer - cleaning: mode imputation}
#---------------------- Mode imputation
# Find mode - look at reference 3
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}

df_cleaning.mode<-df_cleaning
mode.result <- getmode(df_cleaning.mode$BareNuclei)
print(mode.result)


# Fill NAs with mode of 1s
df_cleaning.mode$BareNuclei[is.na(df_cleaning.mode$BareNuclei)] <- mode.result
#check it was imputed correctly
head(df_cleaning.mode,24)
```
### Imputation - regression
```{r reg}
set.seed(123)

newdata<-df_cleaning
missing.index<-which(is.na(newdata$BareNuclei), arr.ind=TRUE)
newdata.1 <- newdata[-missing.index,2:10] # all other predictors data points except for the missing value and response variable
model <- lm(BareNuclei~ClumpThickness+CellSize+CellShape+Adhesion+SingleEpithelialSize+BlandChromatin+NormalNucleoli+Mitoses,data=newdata.1 )
summary(model)
```


```{r breast cancer - cleaning: correlation}
# Correlation viz
df_cleaning_x <- select_if(df_cleaning, is.numeric)
cor(df_cleaning_x)
vis_cor(df_cleaning_x)


# Change Diagnosis from integer to factor
df_cleaning$Diagnosis <- as.factor(df_cleaning$Diagnosis)

pairwise <- ggpairs(df_cleaning, columns = 1:10, ggplot2::aes(colour = Diagnosis), lower=list(combo=wrap("facethist", binwidth=0.5)))
pairwise
```
```{r breast cancer - cleaning: skewness}
################# Skewness
skew(df_cleaning_x)


df_cleaning %>%
  add_prop_miss() %>%
  rpart(prop_miss_all ~ ., data=.) %>%
  prp(type=4, extra = 101, roundint=FALSE, prefix="Prop.Miss = ")
```


## Question 15.1
**Describe a situation or problem from your job, everyday life, current events, etc., for which optimization would be appropriate. What data would you need?**

Optimization is used in descriptive and predictive analytics [1] and there are three parts to it - variables, constraints, and objective function [2]. In November I will be supporting my sister in the NYC Marathon. Applying these three parts to my NYC Marathon cheering strategy, the goal or _objective function_ is to maximize the number of times I can cheer her without missing her. The key _variables_ are the mile markers (where I can stand and cheer). _Constraints_ include traffic, public transportation schedules, and crowd density, which can all affect my adherence to my cheering strategy.



References:

[1] Lecture 1 slides

[2] Lecture 3 slides

[3] R - Mean, Median and Mode - Tutorialspoint. (2019). Tutorialspoint.com. https://www.tutorialspoint.com/r/r_mean_median_mode.htm

[4] APA citation

[5] APA citation
