---
title: "Homework_10"
author: "Claire Kraft"
date: "2024-10-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 14.1

**The breast cancer data set breast-cancer-wisconsin.data.txt from [UCI Repository](http://archive.ics.uci.edu/ml/machine-learning-databases/breast-cancer-wisconsin/) (description at [UCI Dataset Description](http://archive.ics.uci.edu/ml/datasets/Breast+Cancer+Wisconsin+%28Original%29)) has missing values.**

1. **Use the mean/mode imputation method to impute values for the missing data.**

2. **Use regression to impute values for the missing data.**

3. **Use regression with perturbation to impute values for the missing data.**

4. **(Optional) Compare the results and quality of classification models (e.g., SVM, KNN) built using**  
   **(1) the data sets from questions 1, 2, 3;**  
   **(2) the data that remains after data points with missing values are removed;**  
   **(3) the data set when a binary variable is introduced to indicate missing values.**

  
answer
```{r code1}
#code1
```


## Question 15.1
**Describe a situation or problem from your job, everyday life, current events, etc., for which optimization would be appropriate. What data would you need?**



Optimization is used in descriptive and predictive analytics [1] and there are three parts to it - variables, constraints, and objective function [2]. In November I will be supporting my sister in the NYC Marathon. Applying these three parts to my NYC Marathon cheering strategy, the goal or _objective function_ is to maximize the number of times I can cheer her without missing her. The key _variables_ are the mile markers (where I can stand and cheer). _Constraints_ include traffic, public transportation schedules, and crowd density, which can all affect my adherence to my cheering strategy.



References:

[1] Lecture 1 slides

[2] Lecture 3 slides

[3] APA citation

[4] APA citation

[5] APA citation
