---
title: "Homework_4"
author: "Claire Kraft"
date: "2024-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 7.1
**Describe a situation or problem from your job, everyday life, current events, etc., for which exponential smoothing would be appropriate. What data would you need? Would you expect the value of $\alpha A$ (the first smoothing parameter) to be closer to 0 or 1, and why?**

answer

```{r grubb test & whisker plot}
#------------------------- Libraries
#install.packages("outliers")
#install.packages("tidyverse")
library(outliers)
library(ggplot2)

#------------------------- Load and explore data
crime_df <- read.table("C://Users//Clair//OneDrive//Documents//GitHub//omsa//ISYE 6501//Homework 3//uscrime.txt", sep = "\t", header = TRUE)
head(crime_df)
summary(crime_df)

#------------------------- Detect outliers on simulated data
# Perform grubbs test
# look at reference 1
grubbs.test(crime_df$Crime,type=11, opposite = FALSE, two.sided = FALSE) 

#------------------------- Visualize original crime data
# look at reference 2
# Check the column names in your dataset
colnames(crime_df)
# Sample data
data(crime_df)
# Simple boxplot 
ggplot(crime_df, aes(x = "", y = Crime)) + 
    geom_boxplot() + 
    labs(title = "Number of Crimes Per 100,000 People")
```



## Question 7.2
**Using the 20 years of daily high temperature data for Atlanta (July through October) from Question 6.2 (file temps.txt), build and use an exponential smoothing model to help make a judgment of whether the unofficial end of summer has gotten later over the 20 years.  (Part of the point of this assignment is for you to think about how you might use exponential smoothing to answer this question. Feel free to combine it with other models if you’d like to. There’s certainly more than one reasonable approach.)**

**Note: in R, you can use either HoltWinters (simpler to use) or the smooth package’s es function (harder to use, but more general).  If you use es, the Holt-Winters model uses model=”AAM” in the function call (the first and second constants are used “A”dditively, and the third (seasonality) is used “M”ultiplicatively; the documentation doesn’t make that clear).**


answer


References:

[1] source

[2] source

[3] source

[4] source

[5] source
