---
title: "Homework_8"
author: "Claire Kraft"
date: "2024-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 11.1
**Using the crime data set uscrime.txt from Questions 8.2, 9.1, and 10.1, build a regression model using:**
  1. Stepwise regression
  2. Lasso
  3. Elastic net
**For Parts 2 and 3, remember to scale the data first – otherwise, the regression coefficients will be on different scales and the constraint won’t have the desired effect.**

answer
```{r code}
# Read in data
uscrime <- read.table("~/GitHub/omsa/ISYE 6501/Homework 08/uscrime.txt", stringsAsFactors = FALSE, header = TRUE)
head(uscrime)
```


### Question 11.2 Part 2
**For Parts 2 and 3, remember to scale the data first – otherwise, the regression coefficients will be on different scales and the constraint won’t have the desired effect. For Parts 2 and 3, use the glmnet function in R.**

answer
```{r code}
#code
```

### Question 11.2 Part 3
****For Parts 2 and 3, remember to scale the data first – otherwise, the regression coefficients will be on different scales and the constraint won’t have the desired effect. For Parts 2 and 3, use the glmnet function in R.****

answer
```{r code}
#code
```



References:

[1] APA citation

[2] APA citation

[3] APA citation

[4] APA citation

[5] APA citation
