---
title: "Homework_6"
author: "Claire Kraft"
date: "2024-MM-DD"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 9.1
**Using the same crime data set uscrime.txt as in Question 8.2, apply Principal Component Analysis and then create a regression model using the first few principal components. Specify your new model in terms of the original variables (not the principal components), and compare its quality to that of your solution to Question 8.2. You can use the R function prcomp for PCA. (Note that to first scale the data, you can include scale. = TRUE to scale as part of the PCA function. Don’t forget that, to make a prediction for the new city, you’ll need to unscale the coefficients (i.e., do the scaling calculation in reverse)!)**

answer
```{r code}
# Set Up
rm(list = ls())

# Helper 
# Install.packages("GGally") # https://ggobi.github.io/ggally/
library(GGally)
# Install.packages('corrplot') # https://www.rdocumentation.org/packages/corrplot/versions/0.94
library(corrplot)
library(DAAG)

# Read In Data 
uscrime <- read.table("~/GitHub/omsa/ISYE 6501/Homework 06/uscrime.txt", stringsAsFactors = FALSE, header = TRUE)

# Draw Correlations
# Look At Reference 1
crime_data <- cor(uscrime)
corrplot(crime_data, method = 'circle', order = 'AOE')

# Examine Some Of The Strongest Correlations - Choosing These Parameters Based On My Hw 5 Results
# Look At Reference 2
ggpairs(uscrime, columns = c("Ed", "Ineq", "Po1")) 
ggpairs

# PCA
PCA = prcomp(uscrime[,1:15], scale. = TRUE)
summary(PCA)

# Eigenvector Matrix
PCA$rotation  
```



References:

[1] Wei, T., & Simko, V. (2021, November 18). An Introduction to corrplot Package. Cran.r-Project.org. https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html 

[2] APA citation

[3] APA citation

[4] APA citation

[5] APA citation
