---
title: "Homework_2"
author: "Claire Kraft"
date: "2024-09-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 5.1
**Using crime data from the file uscrime.txt (http://www.statsci.org/data/general/uscrime.txt, description at http://www.statsci.org/data/general/uscrime.html), test to see whether there are any outliers in the last column (number of crimes per 100,000 people).  Use the grubbs.test function in the outliers package in R.**

I first ran the grubb test function and chose the second test with this argument ```type=11``` which checks for the opposite extreme ends of the data. It compares the range (difference between the smallest and largest values) to the standard deviation of the data [1]. The p-value is a metric for how valid our observation is. The lower the p value the more statistically significant the observation, the higher the less statistically significant. The p-value is 1 for this grub test. It's best to take this observation with a grain of salt as this is statistically _insignificant_ [3]. 

I then ran the whiskers box plot and there are three outliers [2].

```{r grubb test & whisker plot}
#------------------------- Libraries
#install.packages("outliers")
#install.packages("tidyverse")
library(outliers)
library(ggplot2)

#------------------------- Load and explore data
crime_df <- read.table("C://Users//Clair//OneDrive//Documents//GitHub//omsa//ISYE 6501//Homework 3//uscrime.txt", sep = "\t", header = TRUE)
head(crime_df)
summary(crime_df)

#------------------------- Detect outliers on simulated data

# Perform grubbs test
# look at reference 1
grubbs.test(crime_df$Crime,type=11, opposite = FALSE, two.sided = FALSE) 

#------------------------- Visualize original crime data
# look at reference 2
# Check the column names in your dataset
colnames(crime_df)
# Sample data
data(crime_df)
# Simple boxplot 
ggplot(crime_df, aes(x = "", y = Crime)) + 
    geom_boxplot() + 
    labs(title = "Number of Crimes Per 100,000 People")
```



## Question 6.1
**Describe a situation or problem from your job, everyday life, current events, etc., for which a Change Detection model would be appropriate. Applying the CUSUM technique, how would you choose the critical value and the threshold?**

answer


### Question 6.2.1
**Using July through October daily-high-temperature data for Atlanta for 1996 through 2015, use a CUSUM approach to identify when unofficial summer ends (i.e., when the weather starts cooling off) each year. You can get the data that you need from the file temps.txt or online, for example at http://www.iweathernet.com/atlanta-weather-records or https://www.wunderground.com/history/airport/KFTY/2015/7/1/CustomHistory.html. You can use R if you’d like, but it’s straightforward enough that an Excel spreadsheet can easily do the job too.**

answer


### Question 6.2.2
**Use a CUSUM approach to make a judgment of whether Atlanta’s summer climate has gotten warmer in that time (and if so, when).**

answer


References:

[1] Grubbs, F.E. (1950). Sample Criteria for testing outlying observations. Ann. Math. Stat. 21, 1, 27-58.

[2] Holtz, Y. (n.d.). Basic ggplot2 boxplot. Www.r-Graph-Gallery.com. https://r-graph-gallery.com/262-basic-boxplot-with-ggplot2.html

[3] Simplilearn. (2022, August 30). What Is P-Value in Statistical Hypothesis? | Simplilearn. Simplilearn.com. https://www.simplilearn.com/tutorials/statistics-tutorial/p-value-in-statistics-hypothesis
[4] 
[5] 
